---
title: "Teaching"
page-layout: article
title-block-banner: true
listing: 
  - id: monash-wcd
    template: listing.ejs
    sort: "date"
    contents:
      - "monash-wcd/*.qmd"
  - id: monash-cwd
    template: listing.ejs
    sort: "date"
    contents:
      - "monash-cwd/*.qmd"
  - id: monash-eda
    template: listing.ejs
    sort: "date"
    contents:
      - "monash-eda/*.qmd"
  - id: monash-iml
    template: listing.ejs
    sort: "date"
    contents:
      - "monash-iml/*.qmd"
---

```{r setup, include = FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

```{css, echo = FALSE}
div.callout-important.callout {
  border-left-color: #ad0059;
  background-color: #f5e1eb;
}

h3 {
  color: black!important;
}

.callout-caption-container {
  color: #ad0059;
  padding-bottom: 10px;
}
```




## Teaching


In the past I have taught or developed:

* [**ETC3250/5250 Introduction to Machine Learning**](https://emitanaka.org/iml/) (100%) at Monash University,
* [**ETC5521 Exploratory Data Analysis**](https://eda.numbat.space/) (50%) at Monash University,
* [**ETC5512 Wild Caught Data**](https://wcd.numbat.space/) (25%) at Monash University,
* [**ETC5523 Communicating with Data**](https://cwd.numbat.space/) (100%) at Monash University,
* [**Introduction to R** materials](https://learnr.numbat.space/),
* **STAT5002 Introduction to Statistics** (50%) at The University of Sydney,
* **STAT3012 Applied Linear Models** (100%) at The University of Sydney,
* **STAT906 Experimental Design** (100%) at the University of Wollongong,
* **MATH1013 Mathematical Modelling** (50%) at The University of Sydney,  
* **MATH1005 Statistics** (100%) at The University of Sydney,


and tutored various mathematics and statistics courses.


## Materials

::: {.callout-important icon=false appearance="simple"}

## <i class="fas fa-poll"></i> Do you use my teaching materials?

If yes, it would be awesome if you could fill out [this very short survey](https://docs.google.com/forms/d/e/1FAIpQLSdVnpDphyWnBeMw21Rs8vMySi5RGYN7kQh_VHU6rT83t8CDbA/viewform?usp=sf_link) for me to get statistics of its use!


:::


### Wild-Caught Data

:::{#monash-wcd}
:::

### Communicating with Data

:::{#monash-cwd}
:::

### Exploratory Data Analysis

:::{#monash-eda}
:::

### Introduction to Machine Learning

:::{#monash-iml}
:::

## Workshop


<ul>
```{r, results="asis"}
small <- function(x) paste0("<span style='font-size:0.75rem;line-height:0.9rem;'>", x, "</span>")
dat <- read_csv("../../cv/data/workshops.csv")
dat %>% 
  filter(parent!="no") %>% 
  mutate(text = glue::glue("<li><b>{title}</b> <a href='{url}'><i class='fas fa-globe'></i></a><br> {small(paste0('Last hosted by the <b>', host, '</b> at ', date_start))}</li>")) %>% 
  pull(text) %>% 
  paste(collapse = "\n") %>% 
  cat()
```
<ul>
```{r, results="asis"}
dat %>% 
  filter(group=="SMfOMAB1",
         parent=="no") %>% 
  mutate(text = glue::glue("<li style='font-size:0.85rem;'>{title} <a href='{url}'><i class='fas fa-link'></i></a></li>")) %>% 
  pull(text) %>% 
  paste(collapse = "\n") %>% 
  cat()
```
</ul>
</ul>

