---
title: "Teaching"
page-layout: article
title-block-banner: true
---

```{r setup, include = FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

```{css, echo = FALSE}
div.callout-important.callout {
  border-left-color: #ad0059;
  background-color: #f5e1eb;
}

.callout-caption-container {
  color: #ad0059;
  padding-bottom: 10px;
}
```


::: {.callout-important icon=false appearance="simple"}

## <i class="fas fa-poll"></i> Do you use my teaching materials?

If yes, it would be awesome if you could fill out [this very short survey](https://docs.google.com/forms/d/e/1FAIpQLSdVnpDphyWnBeMw21Rs8vMySi5RGYN7kQh_VHU6rT83t8CDbA/viewform?usp=sf_link) for me to get statistics of its use!


:::


## Teaching

At Monash University, I teach or develop:

  * [**ETC5512 Wild Caught Data**](https://wcd.numbat.space/) (25%)
  * [**ETC5523 Communicating with Data**](https://cwd.numbat.space/) (100%)
  * [**Introduction to R** materials](https://learnr.numbat.space/)

In the past I have taught:

* [**ETC5521 Exploratory Data Analysis**](https://eda.numbat.space/) (50%) at Monash University,
* **STAT5002 Introduction to Statistics** (50%) at The University of Sydney,
* **STAT3012 Applied Linear Models** (75%) at The University of Sydney,
* **STAT906 Experimental Design** (100%) at the University of Wollongong,

and tutored various mathematics and statistics courses at The Univeristy of Sydney including lecturing for *MATH1013 Mathematical Modelling* (50%).


## Workshop

::: {.callout-important icon=false appearance="simple"}

## <i class="fas fa-chalkboard-teacher"></i> Upcoming workshops


* **Data Visualisation with R**, in-person only at Inverloch, Victoria, Mon 28th Nov 9AM-4.30PM AEST [<i class="fas fa-info-circle"></i>](https://aasc2022.netlify.app/news/workshop-datavis/) [<i class="fas fa-user-plus"></i>](https://aasc2022.netlify.app/registration/)
* **Effective Data Visualisation with R**, online Tue 6th Dec 9AM-12PM AEST [<i class="fas fa-info-circle"></i>](https://numbats.github.io/WOMBAT2022/tutorial/emi.html) [<i class="fas fa-user-plus"></i>](https://www.trybooking.com/events/landing?eid=958263&)

*All proceeds go to supporting my research activities.*

:::



<ul>
```{r, results="asis"}
small <- function(x) paste0("<span style='font-size:0.75rem;line-height:0.9rem;'>", x, "</span>")
dat <- read_csv("../../cv/data/workshops.csv")
dat %>% 
  filter(parent!="no") %>% 
  mutate(text = glue::glue("<li><b>{title}</b> <a href='{url}'><i class='fas fa-globe'></i></a><br> {small(paste0('Last hosted by the <b>', host, '</b> at ', date_start))}</li>")) %>% 
  pull(text) %>% 
  paste(collapse = "\n") %>% 
  cat()
```
<ul>
```{r, results="asis"}
dat %>% 
  filter(group=="SMfOMAB1",
         parent=="no") %>% 
  mutate(text = glue::glue("<li style='font-size:0.85rem;'>{title} <a href='{url}'><i class='fas fa-link'></i></a></li>")) %>% 
  pull(text) %>% 
  paste(collapse = "\n") %>% 
  cat()
```
</ul>
</ul>

